### 2024-05-16 Introduction.tex
Adding minor corrections from viva/annotated thesis

The type of tumour and it's impact is determined by the region of brain it affects. There are three main sections of the brain which each control different aspects of human functioning. Some examples are shown in Figure~\ref{fig:anatomybrain} and include the cerebrum, which is the largest part of the brain and controls thinking, learning, problem solving, emotions, speech, reading, writing, and voluntary movement; the cerebellum controls movement, balance, and posture; and the brain stem which connects the brain to the spinal cord and controls breathing, heart rate, and the nerves and muscles used to see, hear, walk, talk, and eat~\citep{Institute2023}.
\begin{figure}[ht]
    \centering
    \includegraphics[width=0.75\textwidth]{brainanatomy.jpg}
    \caption{Depiction of some key regions of the brain reproduced from~\citep{Institute2023}.}
    \label{fig:anatomybrain}
\end{figure}
There are also different types of brain cells in which the tumour can originate. The brain contains many neurones which transport the electrical signals that allow the brain to operate. These are surrounded by glial cells, of which there are different types which each play a role in supporting and protecting the neurones~\citep{TheBrainTumourCharity2023}. The main types are astrocytes, oligodendrocytes and ependymal cells. When the tumour originates in glial cells it is termed a glioma with subcategories corresponding to each type of glial cell. Astrocytomas, originating in astrocytes, are the most common gliomas with the different grades given different names such as Grade 4 astrocytoma named glioblastoma multiforme~\citep{BrainTumourResearch2023}. Similarly, oligodendrogliomas originate from oligodendrocytes, and ependymomas originate from ependymal cells~\citep{TheBrainTumourCharity2023}. %include meningioma lymphoma medulloblastoma adenoma \\

@@

Neurosurgical resection aims for maximal safe resection of the tumour which balances a high extent of resection while preserving functional regions of brain as improved resection is linked to improved patient outcomes~\citep{Chanbour2022}. To this end, a variety of techniques can be employed to improve resection. These include: a surgical microscope which enables clear visualisation of the surgical field, neuronavigation which allows registration of a probe against an MRI scan, awake or asleep sub-corticol mapping to allow identification of eloquent areas of brain, in some cases fluorescence imaging using 5-aminolevulinic acid (5-ALA) or indocyanine green (ICG) can be used for tumour localisation, and microvascular doppler (MVD) or ICG can be used for tumour vessel localisation~\citep{Chanbour2022, Catapano2018}.

@@

\subsection{Current treatments}\label{sec:introaneurysmtreatments}
Treatment of unruptured aneurysm is determined by the risk of rupture, size, and location in the brain. If the aneurysm is considered to have a low risk of rupture, treatment may focus on limiting risk factors~\citep{NationalHealthService2022}. In ruptured aneurysms or those at higher risk of rupturing, however, treatment is primarily surgical in the form of clipping or endovascular procedures depicted in Figure~\ref{fig:aneurysmtreatment}. The former involves the placement of a metal clip over the opening to prevent further blood flow into the aneurysm~\citep{TheBrainFoundation2023}. Endovascular procedures can include coiling to encourage clotting in the aneurysm to prevent further blood flow, which can be combined with further devices including stents to hold the coil in place~\citep{TheBrainFoundation2023}. The final endovascular technique is flow diversion achieved by placing a cylindrical tube (similar to a stent with a tighter weave) in the vessel over the aneurysm opening to allow normal blood flow~\citep{TheBrainFoundation2023}. 
\begin{figure}[h]
    \centering
    \includegraphics[width=0.9\textwidth]{Aneurysm-treatments.png}
    \caption{An illustration of some key surgical treatment methods for brain aneurysms reproduced from~\citep{TheBrainFoundation2023} where the technique depicted by A is coiling, B is clipping, C is coiling combined with a stent, and D is flow diversion.}
    \label{fig:aneurysmtreatment}
\end{figure}

@@

Oxygen is required by all cells in the body to enable release of energy by respiration. This oxygen is transported to cells by haemoglobin, primarily in the red blood cells of blood, which can take the forms of oxyhaemoglobin and deoxyhaemoglobin based on whether the haemoglobin is bound to oxygen or not respectively. These forms of haemoglobin have very different absorbance characteristics as seen in their extinction spectra in Figure~\ref{fig:Haemoglobinext}. 
\begin{figure}[h]
    \centering 
    \includegraphics[width=0.7\textwidth]{Hb_eps.pdf}
    \caption{The extinction coefficients for oxyhaemoglobin and deoxyhaemoglobin reproduced from~\citep{Prahl1998}.}
    \label{fig:Haemoglobinext}
\end{figure}
Oxygen saturation is defined as the proportion of haemoglobin which is oxygenated i.e. $\frac{HbO_2}{HbO_2 + Hb}$ where $Hb$ is deoxyhaemoglobin and $HbO_2$ is oxyhaemoglobin.

Currently, there are no established, intra-operative, techniques to assess $StO_2$ in neurosurgery and so perfusion is often used as an indicator instead with DSA, MVD, and ICG capable of this as detailed in Sections~\ref{sec:introtumourtreatments} and~\ref{sec:introaneurysmtreatments}. Additionally, there has been significant research into using Laser Speckle Contrast Imaging (LSCI) to visualise blood perfusion, however this is not able to distinguish between $Hb$ and $HbO_2$~\citep{Dunn2012, Zhong2021}. Peri-operative neural monitoring can use jugular venous oxygen saturation (JVOS)~\citep{Raith2020} and cerebral oximetry~\citep{Lian2020} to provide indications of the health of the organ. Both techniques usually make use of near infra-red (NIR) spectroscopy to distinguish between $Hb$ and $HbO_2$. These techniques, however, lack specificity so it is possible for local regions of ischaemia to be present even when these values are within the bounds of normality~\citep{Raith2020, Zhong2021}.

%Due to this lack of available techniques, there have been many proposed methods to extract $StO_2$ from tissue diffuse reflectance spectra. These range from the simplest ratiometric two-wavelength methods to more sophisticated analytical models~\citep{MacKenzie2018} or deep-learning based approaches~\citep{Ayala2023}. Using two or three wavelength models can provide good results but require data to be captured at these precise wavelengths, whereas analytical models can be applied to ranges of wavelengths for which they are developed. These analytical models also enable more detailed modelling of tissue including consideration of melanin when imaging through skin. Alternatively, Monte Carlo methods are an established approach to modelling light-tissue interaction for a range of wavelengths. It is, however, computationally challenging to apply in an inverse problem setting making the estimation of optical properties from tissue spectra non trivial on the basis of Monte Carlo simulations alone. To obtain the diffuse reflectance spectra intra-operatively, these techniques primarily use HSI which enable non-contact imaging with minimal change to surgical workflow. 
Due to this lack of available techniques, there have been many proposed methods to extract $StO_2$ from optical measurements of tissues, including from diffuse reflectance spectra. These models relate the extinction coefficients of haemoglobin to the optical properties of tissues to enable $StO_2$ calculation.

@@

Scattering behaviour of particles in biological tissues can be broadly divided into Rayleigh, Mie, and Raman scattering. Commonly, Rayleigh and Mie scattering are defined as the elastic scattering behaviour by particles significantly smaller or particles comparable to or larger than the incident wavelength of light respectively. More precisely, Mie scattering refers to the scattering of spheres of any size with Rayleigh scattering being the Rayleigh limit of Mie scattering for particles significantly smaller than the wavelength of incident light~\citep{Jacques2013}. Raman scattering refers to the inelastic scattering of light, however this effect is typically much smaller and is not easy to detect without long integration times for typical optical measurement systems~\citep{Kong2015}. Within the common measurement range of 400-1000nm, modelling Mie scattering alone is sufficient, however beyond this range a combination of Rayleigh and Mie scattering may be required to provide a good model of the scattering behaviour~\citep{Jacques2013}. This Rayleigh component at extreme wavelengths can be more significant for tissues with higher levels of collagen, whereas Mie theory continues to provide a good model of scattering for tissues where the predominant scatterer is fat~\citep{Jacques2013}. By using Mie theory, each of the scattering coefficients can be calculated using the diameter and volume fraction of the scattering particles for each wavelength of light~\citep{Hergert2012}. This information, however, is usually unknown for biological tissue and so the Mie scattering is often simplified to Equation~\eqref{eq:Mieintro} where the reduced scattering at 500nm is taken as a reference point and scattering coefficients $a$ and $b$ vary across tissues~\citep{Jacques2013}. 
\begin{equation}
    \mu_s' = a(\frac{\lambda}{500nm})^{-b}
    \label{eq:Mieintro}
\end{equation}
By considering these optical properties of tissues, a variety of methods to extract $StO_2$ from measurements of tissue have been developed.

@@

The most simplistic method of $StO_2$ estimation was the two-wavelength approach based on a pure Beer-Lambert model of tissue. This method assumes only oxygenated and deoxygenated haemoglobin ($HbO_2$ and $Hb$) as chromophores in tissue. 
\begin{equation}
    \mu_a = c_{Hb}\epsilon_{Hb} + c_{HbO_2}\epsilon_{HbO_2}
\end{equation}
In this method measurements of tissue are acquired at two wavelengths, one of which is an isosbestic point for the haemoglobin chromophore where the oxygenated and deoxygenated extinction coefficients are equal so the measurement is invariant to $StO_2$~\citep{MacKenzie2018}. This approach required significant calibration to account for scattering in tissues~\citep{MacKenzie2018}. Pulse oximetry is based on this, however the pulsatile nature of arterial blood allows the background non-pulsatile absorbance to be removed meaning that arterial $SpO_2$ can be calculated for the pulsatile portion of the signal only~\citep{MacKenzie2018}. This, however, is still confounded by the presence of other significant absorbers in high concentrations such as melanin or by low blood flow due to low temperature~\citep{Bangash2022, MacKenzie2018}. This allowed development of the three-wavelength approach. In this case, two isosbestic wavelengths are measured to calculate a wavelength-independent scattering coefficient with a third used for $StO_2$ extraction~\citep{MacKenzie2018}. Whilst this provides advantages over the two-wavelength approach, it has largely been superseded by hyperspectral approaches due to the limited number of appropriate wavelengths that can be used which satisfy the approximation of constant scattering and by the continued requirement for calibration with samples of known $StO_2$~\citep{MacKenzie2018}. By using ranges of wavelengths and investigating the derivative of the spectrum a hyperspectral alternative has been proposed, however this also assumes wavelength-independent scattering and requires significant calibration~\citep{Holmer2018}. To advance this field, Monte Carlo simulations were used to probabilistically predict the spectra that could be obtained from a sample with known optical properties~\citep{Wang1995}. These have been used to improve approaches based on the Beer-Lambert models by including differential path length factors which are fitted to simulated data~\citep{Clancy2020}.

@@

\subsection{Optical coherence tomography}
Optical Coherence Tomography (OCT) is a non-contact 3D imaging modality which provides both spatial and depth information of the tissue imaged. The technique utilises a reference beam and a sample beam; the backscattered light from the sample beam forms an interference pattern with the reference beam which is measured. Traditionally, the mirror used to produce the reference beam is adjusted until interference is maximised, and at this point the distance between the positions of the mirror and sample provides information on the depth of the scatterer. To increase efficiency, the mirror location is fixed and Fourier Transforms are used to analyse the returned interference and provide the depth of the scattering structures at the spatial location measured~\citep{Fercher2003}. By scanning the focussed beam across the sample tissue, a 3D image with both spatial and depth information can be constructed~\citep{Fercher2003}. Traditionally OCT utilises near-infrared wavelengths (NIR) due to the deeper penetration into biological tissue, however recent advances have shown that using visible wavelengths allows oxygen saturation of vessels to be computed~\citep{Shu2017}. It has also been shown that tumour margins can be highly accurately delineated from this type of data using automated algorithms~\citep{Sunny2019}. 
%Whilst this provides spatially resolved oxygenation information of major vessels, significant attenuation is required for this calculation limiting its application to tissue directly~\citep{Shu2017}. 
This technique allows spatially-resolved, intra-operative, oxygen saturation measurement of vessels and tumour margin delineation using probes that integrate easily into the surgical workflow~\citep{Jansen2018}. Despite these key advantages of OCT, it is also associated with some significant limitations. While OCT can be adapted to measuring oxygen saturation of vessels, significant attenuation of the sample beam is required for this calculation which limits its application to tissue directly~\citep{Shu2017}. Additionally wide-field OCT obtained by scanning the focussed sample beam over the imaging area can provide good spatially-resolved information, however this is subject to motion artefacts of the tissue under examination and limits real-time wide-field implementation~\citep{Yu2015}. These limitations are significant in the context of neurosurgery where local tissue ischaemia is of significant concern and wide-field imaging provides important information for decision-making.

@@

\subsection{Photoacoustic imaging}
Photoacoustic imaging is a hybrid imaging modality which optically illuminates a sample, the sample then converts the optical energy to thermal energy which generates acoustic waves that are detected. Scattering of sound waves is much lower than of light in tissues so enabling deeper imaging, however increasing the depth of imaging compromises the spatial resolution~\citep{Assi2023}. By utilising different wavelengths of light, spatially-resolved $StO_2$ can be determined. This increase in the number of wavelengths used for imaging results in slower acquisition times due to the increased scanning which can result in motion artefacts in the acquired image and limits real-time implementation~\citep{Assi2023, Attia2019}. By considering different biological or exogeneous chromophores, tumour margins can also be detected with this technique which has been primarily explored for breast cancers and melanomas~\citep{Assi2023, Attia2019, Taylor-Williams2022}. Quantitative $StO_2$ or chromophore extraction is challenging at depth due to the absorption of the incident light by chromophores in the sample, termed "spectral colouring" whose corrections often break down in biological settings~\citep{Assi2023, Taylor-Williams2022}. While some photoacoustic devices have received CE marking, many devices use high power lasers for illumination, however these come with associated risks~\citep{Assi2023}. These risks are mitigated by conforming to existing laser safety standards for eye and skin exposure, however this does not account for internal tissues, including brain tissues, that would be exposed in intra-operative use~\citep{Assi2023}. Finally, there are many different possible imaging configurations that can be used for this technology which range in ergonomics~\citep{Attia2019} and can require highly specialised training for operation~\citep{Assi2023}. Whilst photoacoustic imaging has been shown to be applicable clinically for $StO_2$ extraction, this does not provide quantitative or real-time results and many imaging configurations present challenges for intra-operative integration.

\subsection{Probe-based diffuse reflectance spectroscopy}
When light is incident on a surface it can either be directly reflected in a specular (mirror-like) manner, or it can be internally scattered or absorbed before leaving the sample diffusely, as seen in Figure~\ref{fig:diffuseR}. Specularly scattered light can carry topological information of the sample, this is examined using laser speckle spectroscopy which can investigate blood flow~\citep{Dunn2012}, however diffusely scattered light is linked to the absorption and scattering properties of the sample. 
\begin{figure}[h]
    \centering
    \includegraphics[width=0.4\textwidth]{Tissue_figure.pdf}
    \caption{Schematic of specularly and diffusely reflected light.}
    \label{fig:diffuseR}
\end{figure}
Diffuse reflectance spectroscopy measures this diffusely scattered light from a sample in order to gain information on the chemical composition of the sample by investigating the absorbance and scattering properties of the sample. Whilst this work focuses on HSI as a method of diffuse reflectance spectroscopy, there are many fibre-optic probe-based systems for these measurements. These systems provide highly accurate diffuse reflectance spectra in a point-based measurement. Probe-based methods are able to provide high spectral resolution measurements with short acquisition times, whereas HSI usually requires longer acquisition periods for similar spectral quality~\citep{Dinish2017}. This high spectral resolution of probe-based systems comes at the cost of spatial resolution. Whilst measurements from multiple probe locations can be stitched together~\citep{Thrapp2020}, wide-field imaging is not possible~\citep{Nishidate2015}. This technique traditionally involves contact between the probe and the sample, however this contact can alter optical properties of the tissue sample~\citep{Miller2017}. To mitigate this, flexible probes have been proposed~\citep{Miller2017} or non-contact solutions have been developed which make use of lenses and remove the impact of specular reflections using polarising filters~\citep{Bish2011, Zhu2012}. %None of these measurement configurations, however, provide wide-field information. 
Diffuse reflectance spectroscopy has shown potential for tissue differentiation~\citep{Skyrman2022} and $StO_2$ measurement~\citep{Fredriksson2020} using methods that are also applied to HSI and are discussed more thoroughly in Sections~\ref{sec:introoptics} and~\ref{sec:introHSI}. These techniques are able to provide real-time, intra-operative, highly accurate measurements which show promise for tissue differentiation and $StO_2$ measurement, however they do not provide wide-field imaging and non-contact configurations introduce significantly more bulk to the system. For this reason HSI is investigated to overcome these limitations.

>>>>

The type of tumour and it's impact is determined by the region of brain it affects. There are three main sections of the brain which each control different aspects of human functioning. Some examples are shown in Figure~\ref{fig:anatomybrain} and include the cerebrum, which is the largest part of the brain and controls thinking, learning, problem solving, emotions, speech, reading, writing, and voluntary movement; the cerebellum controls movement, balance, and posture; and the brain stem which connects the brain to the spinal cord and controls breathing, heart rate, and the nerves and muscles used to see, hear, walk, talk, and eat~\citep{Institute2023}.
\begin{figure}[ht]
    \centering
    \includegraphics[width=0.75\textwidth]{brainanatomy.png}
    \caption{Depiction of some key regions of the brain reproduced from~\citep{NursingHero2024}.}
    \label{fig:anatomybrain}
\end{figure}
There are also different types of brain cells in which the tumour can originate. The brain contains many neurones which transport the electrical signals that allow the brain to operate. These are surrounded by glial cells, of which there are different types which each play a role in supporting and protecting the neurones~\citep{TheBrainTumourCharity2023}. The main types are astrocytes, oligodendrocytes and ependymal cells. When the tumour originates in glial cells it is termed a glioma with subcategories corresponding to each type of glial cell. Astrocytomas, originating in astrocytes, are the most common gliomas with the different grades given different names such as Grade 4 astrocytoma named glioblastoma multiforme~\citep{BrainTumourResearch2023}. Similarly, oligodendrogliomas originate from oligodendrocytes, and ependymomas originate from ependymal cells~\citep{TheBrainTumourCharity2023}. %include meningioma lymphoma medulloblastoma adenoma \\

@@

Neurosurgical resection aims for maximal safe resection of the tumour which balances a high extent of resection while preserving functional regions of brain as improved resection is linked to improved patient outcomes~\citep{Chanbour2022}. \textcolor{red}{Tumour boundaries are assessed visually intraoperatively and so high quality visualisation is required for accurate resection.~\citep{Chanbour2022}} \annref{4} To this end, a variety of techniques can be employed to improve resection. These include: a surgical microscope which enables clear visualisation of the surgical field, neuronavigation which allows registration of a probe against an MRI scan, awake or asleep sub-corticol mapping to allow identification of eloquent areas of brain, in some cases fluorescence imaging using 5-aminolevulinic acid (5-ALA) or indocyanine green (ICG) can be used for tumour localisation, and microvascular doppler (MVD) or ICG can be used for tumour vessel localisation~\citep{Chanbour2022, Catapano2018}.

@@

\subsection{Current treatments}\label{sec:introaneurysmtreatments}
Treatment of unruptured aneurysm is determined by the risk of rupture, size, and location in the brain. If the aneurysm is considered to have a low risk of rupture, treatment may focus on limiting risk factors~\citep{NationalHealthService2022}. In ruptured aneurysms or those at higher risk of rupturing, however, treatment is primarily surgical in the form of clipping or endovascular procedures depicted in Figure~\ref{fig:aneurysmtreatment}. The former involves the placement of a metal clip over the opening to prevent further blood flow into the aneurysm~\citep{TheBrainFoundation2023}. Endovascular procedures can include coiling to encourage clotting in the aneurysm to prevent further blood flow, which can be combined with further devices including stents to hold the coil in place~\citep{TheBrainFoundation2023}. The final endovascular technique is flow diversion achieved by placing a cylindrical tube (similar to a stent with a tighter weave) in the vessel over the aneurysm opening to allow normal blood flow~\citep{TheBrainFoundation2023}. 
\begin{figure}[h]
    \centering
    \includegraphics[width=0.9\textwidth]{Aneurysm-treatments.png}
    \caption{An illustration of some key surgical treatment methods for brain aneurysms reproduced with permission from~\citep{Perrone2015} where the technique depicted by A is coiling, B is clipping, C is coiling combined with a stent, and D is flow diversion.}
    \label{fig:aneurysmtreatment}
\end{figure}

@@

Oxygen is required by all cells in the body to enable release of energy by respiration. This oxygen is transported to cells by haemoglobin, primarily in the red blood cells of blood, which can take the forms of oxyhaemoglobin and deoxyhaemoglobin based on whether the haemoglobin is bound to oxygen or not respectively. These forms of haemoglobin have very different absorbance characteristics as seen in their extinction spectra in Figure~\ref{fig:Haemoglobinext}. 
\begin{figure}[h]
    \centering 
    \includegraphics[width=0.7\textwidth]{Hb_eps.pdf}
    \annref{6}
    \caption{The extinction coefficients for oxyhaemoglobin and deoxyhaemoglobin \textcolor{red}{in the visible wavelength range (400-700~nm)} reproduced from~\citep{Prahl1998}.}
    \label{fig:Haemoglobinext}
\end{figure}
Oxygen saturation is defined as the proportion of haemoglobin which is oxygenated i.e. $\frac{HbO_2}{HbO_2 + Hb}$\textcolor{red}{$\times 100\%$} \annref{5} where $Hb$ is deoxyhaemoglobin and $HbO_2$ is oxyhaemoglobin.

Currently, there are no established, intra-operative, techniques to assess $StO_2$ in neurosurgery and so perfusion is often used as an indicator instead with DSA, MVD, and ICG capable of this as detailed in Sections~\ref{sec:introtumourtreatments} and~\ref{sec:introaneurysmtreatments}. Additionally, there has been significant research into using Laser Speckle Contrast Imaging (LSCI) to visualise blood perfusion, however this is not able to distinguish between $Hb$ and $HbO_2$~\citep{Dunn2012, Zhong2021}. Peri-operative neural monitoring can use jugular venous oxygen saturation (JVOS)~\citep{Raith2020} and cerebral oximetry~\citep{Lian2020} to provide indications of the health of the organ. Both techniques usually make use of near infra-red (NIR) spectroscopy to distinguish between $Hb$ and $HbO_2$. These techniques, however, \annref{7} \textcolor{red}{cannot provide spatially resolved or tissue specific measurements} so it is possible for local regions of ischaemia to be present even when these values are within the bounds of normality~\citep{Raith2020, Zhong2021}.

%Due to this lack of available techniques, there have been many proposed methods to extract $StO_2$ from tissue diffuse reflectance spectra. These range from the simplest ratiometric two-wavelength methods to more sophisticated analytical models~\citep{MacKenzie2018} or deep-learning based approaches~\citep{Ayala2023}. Using two or three wavelength models can provide good results but require data to be captured at these precise wavelengths, whereas analytical models can be applied to ranges of wavelengths for which they are developed. These analytical models also enable more detailed modelling of tissue including consideration of melanin when imaging through skin. Alternatively, Monte Carlo methods are an established approach to modelling light-tissue interaction for a range of wavelengths. It is, however, computationally challenging to apply in an inverse problem setting making the estimation of optical properties from tissue spectra non trivial on the basis of Monte Carlo simulations alone. To obtain the diffuse reflectance spectra intra-operatively, these techniques primarily use HSI which enable non-contact imaging with minimal change to surgical workflow. 
Due to this lack of available techniques, there have been many proposed methods to extract $StO_2$ from optical measurements of tissues, including from diffuse reflectance spectra. These models relate the extinction coefficients of haemoglobin to the optical properties of tissues to \annref{8}\textcolor{red}{either enable $StO_2$ calculation directly or to extract absolute or relative changes in $StO_2$}.

@@

Scattering behaviour of particles in biological tissues can be broadly divided into Rayleigh, Mie, and Raman scattering. Commonly, Rayleigh and Mie scattering are defined as the elastic scattering behaviour by particles significantly smaller or particles comparable to or larger than the incident wavelength of light respectively. More precisely, Mie scattering refers to the scattering of spheres of any size with Rayleigh scattering being the Rayleigh limit of Mie scattering for particles significantly smaller than the wavelength of incident light~\citep{Jacques2013}. Raman scattering refers to the inelastic scattering of light, however this effect is typically much smaller and is not easy to detect without long integration times for typical optical measurement systems~\citep{Kong2015}. \annref{9} \textcolor{red}{Commonly measurements are taken within the wavelength range of 400-1000~nm~\citep{Jacques2013}. Within this range}, modelling Mie scattering alone is sufficient, however beyond this range a combination of Rayleigh and Mie scattering may be required to provide a good model of the scattering behaviour~\citep{Jacques2013}. This Rayleigh component at extreme wavelengths can be more significant for tissues with higher levels of collagen, whereas Mie theory continues to provide a good model of scattering for tissues where the predominant scatterer is fat~\citep{Jacques2013}. By using Mie theory, each of the scattering coefficients can be calculated using the diameter and volume fraction of the scattering particles for each wavelength of light~\citep{Hergert2012}. This information, however, is usually unknown for biological tissue and so the Mie scattering is often simplified to Equation~\eqref{eq:Mieintro} where the reduced scattering at 500nm is taken as a reference point and scattering coefficients $a$ and $b$ vary across tissues~\citep{Jacques2013}. 
\begin{equation}
    \mu_s' = a(\frac{\lambda}{500nm})^{-b}
    \label{eq:Mieintro}
\end{equation}
By considering these optical properties of tissues, a variety of methods to extract $StO_2$ from measurements of tissue have been developed.

@@

The most simplistic method of $StO_2$ estimation was the two-wavelength approach based on a pure Beer-Lambert model of tissue. This method assumes only oxygenated and deoxygenated haemoglobin ($HbO_2$ and $Hb$) as chromophores in tissue. 
\begin{equation}
    \mu_a = c_{Hb}\epsilon_{Hb} + c_{HbO_2}\epsilon_{HbO_2}
\end{equation}
In this method measurements of tissue are acquired at two wavelengths, one of which is an isosbestic point for the haemoglobin chromophore where the oxygenated and deoxygenated extinction coefficients are equal so the measurement is invariant to $StO_2$ \annref{10}\textcolor{red}{, $\mu_{a, iso} = \epsilon_{Hb, iso}(c_{Hb} + c_{HbO_2})$}~\citep{MacKenzie2018}. \annref{10}\textcolor{red}{By taking the ratio of these two points, the $StO_2$ can be calculated.} This approach required significant calibration to account for scattering in tissues~\citep{MacKenzie2018}. \annref{11} \textcolor{red}{Some p}ulse oximetry is based on this, however the pulsatile nature of arterial blood allows the background non-pulsatile absorbance to be removed meaning that arterial $SpO_2$ can be calculated for the pulsatile portion of the signal only~\citep{MacKenzie2018}. This, however, is still confounded by the presence of other significant absorbers in high concentrations such as melanin or by low blood flow due to low temperature~\citep{Bangash2022, MacKenzie2018}. This allowed development of the three-wavelength approach. In this case, two isosbestic wavelengths are measured to calculate a wavelength-independent scattering coefficient with a third used for $StO_2$ extraction~\citep{MacKenzie2018}. Whilst this provides advantages over the two-wavelength approach, it has largely been superseded by hyperspectral approaches due to the limited number of appropriate wavelengths that can be used which satisfy the approximation of constant scattering and by the continued requirement for calibration with samples of known $StO_2$~\citep{MacKenzie2018}. \annref{12} \textcolor{red}{Increasing the number of wavelengths considered is also advantageous as this likely reduces the impact of spectral noise on the $StO_2$ extracted.} By using ranges of wavelengths and investigating the derivative of the spectrum \annref{13} \textcolor{red}{with respect to wavelength,} a hyperspectral alternative has been proposed, however this also assumes wavelength-independent scattering and requires significant calibration~\citep{Holmer2018}. To advance this field, Monte Carlo simulations were used to probabilistically predict the spectra that could be obtained from a sample with known optical properties~\citep{Wang1995}. These have been used to improve approaches based on the Beer-Lambert models by including differential path length factors which are fitted to simulated data~\citep{Clancy2020}.

@@

\subsection{Optical coherence tomography}
Optical Coherence Tomography (OCT) is a non-contact 3D imaging modality which provides both spatial and depth information of the tissue imaged. The technique utilises a reference beam and a sample beam; the backscattered light from the sample beam forms an interference pattern with the reference beam which is measured. Traditionally, the mirror used to produce the reference beam is adjusted until interference is maximised, and at this point the distance between the positions of the mirror and sample provides information on the depth of the scatterer. To increase efficiency, the mirror location is fixed and Fourier Transforms are used to analyse the returned interference and provide the depth of the scattering structures at the spatial location measured~\citep{Fercher2003}. By scanning the focussed beam across the sample tissue, a 3D image with both spatial and depth information can be constructed~\citep{Fercher2003}. Traditionally OCT utilises near-infrared wavelengths (NIR) due to the deeper penetration into biological tissue, however recent advances have shown that using visible wavelengths allows oxygen saturation of vessels to be computed~\citep{Shu2017}. It has also been shown that tumour margins can be highly accurately delineated from this type of data using \annref{14}\textcolor{red}{machine learning} algorithms~\citep{Sunny2019}. 
%Whilst this provides spatially resolved oxygenation information of major vessels, significant attenuation is required for this calculation limiting its application to tissue directly~\citep{Shu2017}. 
This technique allows spatially-resolved, intra-operative, oxygen saturation measurement of vessels and tumour margin delineation using probes that integrate easily into the surgical workflow~\citep{Jansen2018}. Despite these key advantages of OCT, it is also associated with some significant limitations. While OCT can be adapted to measuring oxygen saturation of vessels, significant attenuation of the sample beam is required for this calculation which limits its application to tissue directly~\citep{Shu2017}. Additionally wide-field OCT obtained by scanning the focussed sample beam over the imaging area can provide good spatially-resolved information, however this is subject to motion artefacts of the tissue under examination and limits real-time wide-field implementation~\citep{Yu2015}. These limitations are significant in the context of neurosurgery where local tissue ischaemia is of significant concern and wide-field imaging provides important information for decision-making.

@@

\subsection{Photoacoustic imaging}
Photoacoustic imaging is a hybrid imaging modality which optically illuminates a sample, the sample \annref{15}\textcolor{red}{absorbs this energy and} then converts the optical energy to thermal energy which generates acoustic waves that are detected. Scattering of sound waves is much lower than of light in tissues so enabling deeper imaging, however increasing the depth of imaging compromises the spatial resolution~\citep{Assi2023}. By utilising different wavelengths of light, spatially-resolved $StO_2$ can be determined. This increase in the number of wavelengths used for imaging results in slower acquisition times due to the increased scanning which can result in motion artefacts in the acquired image and limits real-time implementation~\citep{Assi2023, Attia2019}. By considering different biological or exogeneous chromophores, tumour margins can also be detected with this technique which has been primarily explored for breast cancers and melanomas~\citep{Assi2023, Attia2019, Taylor-Williams2022}. Quantitative $StO_2$ or chromophore extraction is challenging at depth due to the absorption of the incident light by chromophores in the sample, termed "spectral colouring" whose corrections often break down in biological settings~\citep{Assi2023, Taylor-Williams2022}. While some photoacoustic devices have received CE marking, many devices use high power lasers for illumination, however these come with associated risks~\citep{Assi2023}. These risks are mitigated by conforming to existing laser safety standards for eye and skin exposure, however this does not account for internal tissues, including brain tissues, that would be exposed in intra-operative use~\citep{Assi2023}. Finally, there are many different possible imaging configurations that can be used for this technology which range in ergonomics~\citep{Attia2019} and can require highly specialised training for operation~\citep{Assi2023}. Whilst photoacoustic imaging has been shown to be applicable clinically for $StO_2$ extraction, this does not provide \annref{16} real-time results and many imaging configurations present challenges for intra-operative integration.

\subsection{Probe-based diffuse reflectance spectroscopy}
When light is incident on a surface it can either be directly reflected in a specular (mirror-like) manner, or it can be internally scattered or absorbed before leaving the sample diffusely, as seen in Figure~\ref{fig:diffuseR}. Specularly scattered light can carry topological information of the sample, this is examined using laser speckle spectroscopy which can investigate blood flow~\citep{Dunn2012}, however diffusely scattered light is linked to the absorption and scattering properties of the sample. 
\begin{figure}[h]
    \centering
    \includegraphics[width=0.4\textwidth]{Tissue_figure.pdf}
    \caption{Schematic of specularly and diffusely reflected light.}
    \label{fig:diffuseR}
\end{figure}
Diffuse reflectance spectroscopy measures this diffusely scattered light from a sample in order to gain information on the chemical composition of the sample by investigating the absorbance and scattering properties of the sample. Whilst this work focuses on HSI as a method of diffuse reflectance spectroscopy, there are many fibre-optic probe-based systems for these measurements. These systems provide highly accurate diffuse reflectance spectra in a point-based measurement. Probe-based methods are able to provide high spectral resolution measurements with short acquisition times, whereas HSI usually requires longer acquisition periods for similar spectral quality~\citep{Dinish2017}. This high spectral resolution of probe-based systems comes at the cost of spatial resolution. Whilst measurements from multiple probe locations can be stitched together~\citep{Thrapp2020}, wide-field imaging is not possible~\citep{Nishidate2015}. This technique traditionally involves contact between the probe and the sample, however this contact can alter optical properties of the tissue sample~\citep{Miller2017}. To mitigate this, flexible probes have been proposed~\citep{Miller2017} or non-contact solutions have been developed which make use of lenses and remove the impact of specular reflections using polarising filters\annref{17}\textcolor{red}{, however this introduces more bulk to the system}~\citep{Bish2011, Zhu2012}. %None of these measurement configurations, however, provide wide-field information. 
Diffuse reflectance spectroscopy has shown potential for tissue differentiation~\citep{Skyrman2022} and $StO_2$ measurement~\citep{Fredriksson2020} using methods that are also applied to HSI and are discussed more thoroughly in Sections~\ref{sec:introoptics} and~\ref{sec:introHSI}. These techniques are able to provide real-time, intra-operative, highly accurate measurements which show promise for tissue differentiation and $StO_2$ measurement, however they do not provide wide-field imaging and non-contact configurations introduce significantly more bulk to the system. For this reason HSI is investigated to overcome these limitations.

### 2024-05-16 List_of_Publications.tex
Adding minor corrections from viva/annotated thesis

\item \bibentry{Bahl2023a}
% \input{PhD-Thesis-LOP_1.bib}
%Update when 1-layer published
%Add in 2-layer paper when preprint or published
\end{itemize}

\section*{First-author articles in international conferences with published proceedings}
%Add tissue modelling reference when proceedings published

>>>>

\textcolor{red}{\item \bibentry{Bahl2023a}} \annref{3} 
% \item Anisha Bahl, Silvere Segaud, Yijing Xie, Jonathan Shapey, Mads Bergholt and Tom Vercauteren. \textit{A comparative study of analytical models of diffuse reflectance in homogeneous biological tissues: Gelatin based phantoms and monte carlo experiments.} Under review.

% \item \bibentry{Bahl2024}

\item Anisha Bahl, Jonathan Shapey, Mads Bergholt and Tom Vercauteren. \textcolor{red}{(2024).} \textit{Evaluation of a two-layer skin tissue model for diffuse reflectance:
A simulated and experimental data study.} Under \textcolor{red}{review}. \annref{3}
% \item \bibentry{Bahl2023a}
% \input{PhD-Thesis-LOP_1.bib}
%Update when 1-layer published
%Add in 2-layer paper when preprint or published
\end{itemize}

\section*{First-author articles in international conferences with published proceedings}%Add tissue modelling reference via bibtex when proceedings published 
\begin{itemize}
\item \textcolor{red}{\bibentry{Bahl2024a}} \annref{3} 
% \item \bibentry{Bahl2024a}
% \item Anisha Bahl, Silvere Segaud, Yijing Xie, Jonathan Shapey, Mads Bergholt and Tom Vercauteren. \textit{A comparative study of analytical tissue reflectance models
% using simulated and experimental data.} Awaiting publication.
\end{itemize}
